ABCD_OHBM_PFN_Analyses

### 2021-11-24 ABCD_OHBM_PFN_Analyses.Rmd
Added correlation matrix

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width=250)
library(tidyr)
library(dplyr)
library(ggplot2)
library(lsr)
library(lme4)
library(lmerTest)
```

@@

# calculate a total FPN score
combined.data.noOutliers$totalFPN <- rowSums(data.frame(as.numeric(combined.data.noOutliers$df4),as.numeric(combined.data.noOutliers$df16),as.numeric(combined.data.noOutliers$df18)))

# separate train and test sets
trainIDs <- read.csv("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/train.csv")
train.data <- merge(combined.data.noOutliers,trainIDs,by="subjectkey") # n=2,149
test.data <- combined.data.noOutliers[!is.element(combined.data.noOutliers$subjectkey,trainIDs$subjectkey),] #n=2,424
```

# Run some models!
```{r}

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width=250)
library(tidyr)
library(dplyr)
library(ggplot2)
library(lsr)
library(lme4)
library(lmerTest)
library(mediation)
library(ggcorrplot)
library(corrplot)
library(RColorBrewer)
```

@@

# calculate a total FPN score
abcd.data$totalFPN <- rowSums(data.frame(as.numeric(abcd.data$df4),as.numeric(abcd.data$df16),as.numeric(abcd.data$df18)))

# separate train and test sets
trainIDs <- read.csv("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/train.csv")
train.data <- merge(abcd.data,trainIDs,by="subjectkey") # n=2,148
test.data <- abcd.data[!is.element(abcd.data$subjectkey,trainIDs$subjectkey),] #n=2,151
```

# Run some models!
```{r}
# Network 17 is specifically associated with PC1 and not PC2 or PC3
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+df18,data=train.data)) #***
summary(lmer(thompson_PC2~interview_age+sex+(1|rel_family_id)+df18,data=train.data))
summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+df18,data=train.data))

# Network 15 is specifically associated with PC1 and PC3 but not PC2
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+df16,data=train.data)) #**
summary(lmer(thompson_PC2~interview_age+sex+(1|rel_family_id)+df16,data=train.data))
summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+df16,data=train.data)) #**

# Network 3 is not associated with any of the PCs
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+df4,data=train.data)) 
summary(lmer(thompson_PC2~interview_age+sex+(1|rel_family_id)+df4,data=train.data))
summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+df4,data=train.data))

# We already know that trauma and physical neglect are associated with impairment in PC1
# Physical neglect is also associated with impairment in PC3
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+trauma.binary,data=train.data)) #**
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data)) #***
summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data)) #***

# Are any subtypes of ACEs associated with PFN topography?
summary(lmer(totalFPN~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data)) # ns
summary(lmer(totalFPN~interview_age+sex+(1|rel_family_id)+acceptance.binary,data=train.data)) # ns
summary(lmer(totalFPN~interview_age+sex+(1|rel_family_id)+abuse.binary,data=train.data)) # ns
summary(lmer(totalFPN~interview_age+sex+(1|rel_family_id)+trauma.binary,data=train.data)) # ns

summary(lmer(df18~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data)) # ns
summary(lmer(df18~interview_age+sex+(1|rel_family_id)+acceptance.binary,data=train.data)) # ns
summary(lmer(df18~interview_age+sex+(1|rel_family_id)+abuse.binary,data=train.data)) # ns
summary(lmer(df18~interview_age+sex+(1|rel_family_id)+trauma.binary,data=train.data)) # ns

summary(lmer(df16~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data)) # ns (0.056)
summary(lmer(df16~interview_age+sex+(1|rel_family_id)+acceptance.binary,data=train.data)) # ns
summary(lmer(df16~interview_age+sex+(1|rel_family_id)+abuse.binary,data=train.data)) # ns
summary(lmer(df16~interview_age+sex+(1|rel_family_id)+trauma.binary,data=train.data)) # ns

summary(lmer(df4~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data)) # ns
summary(lmer(df4~interview_age+sex+(1|rel_family_id)+acceptance.binary,data=train.data)) # ns
summary(lmer(df4~interview_age+sex+(1|rel_family_id)+abuse.binary,data=train.data)) # ns
summary(lmer(df4~interview_age+sex+(1|rel_family_id)+trauma.binary,data=train.data)) # ns

@@

# Mediation analyses -- these are significant!
```{r}
# Mediation: trauma --> Network 17 --> PC1 (path C does decrease) *
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+trauma.binary,data=train.data))
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+trauma.binary+df18,data=train.data))

bs_indir = c()
for (i in 1:1000){
  thisSample <- sample(c(1:length(train.data$subjectkey)),length(train.data$subjectkey),replace=TRUE)
  pathc<-lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+trauma.binary,data=train.data[thisSample,])
  pathcp<-lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+trauma.binary+df18,data=train.data[thisSample,])
  indir<-fixef(pathc)[4]-fixef(pathcp)[5]
  bs_indir <- c(bs_indir,indir)
}
t.test(bs_indir,conf.level=0.95)

# Mediation: trauma --> Network 15 --> PC1 (path C does not decrease)
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+trauma.binary,data=train.data))
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+trauma.binary+df16,data=train.data))

# Mediation: physical neglect --> Network 17 --> PC1 (path C does decrease) *
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data))
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+monitoring.binary+df18,data=train.data))

bs_indir = c()
for (i in 1:1000){
  thisSample <- sample(c(1:length(train.data$subjectkey)),length(train.data$subjectkey),replace=TRUE)
  pathc<-lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data[thisSample,])
  pathcp<-lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+monitoring.binary+df18,data=train.data[thisSample,])
  indir<-fixef(pathc)[4]-fixef(pathcp)[5]
  bs_indir <- c(bs_indir,indir)
}
t.test(bs_indir,conf.level=0.95)

# Mediation: physical neglect --> Network 15 --> PC1 (path C does not decrease)
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data))
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+monitoring.binary+df16,data=train.data))

# Mediation: physical neglect --> Network 17 --> PC3 (path C does decrease) *
summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data))
summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+monitoring.binary+df18,data=train.data))

bs_indir = c()
for (i in 1:1000){
  thisSample <- sample(c(1:length(train.data$subjectkey)),length(train.data$subjectkey),replace=TRUE)
  pathc<-lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data[thisSample,])
  pathcp<-lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+monitoring.binary+df18,data=train.data[thisSample,])
  indir<-fixef(pathc)[4]-fixef(pathcp)[5]
  bs_indir <- c(bs_indir,indir)
}
t.test(bs_indir,conf.level=0.95)

# Mediation: physical neglect --> Network 15 --> PC3 (path C does not decrease)
summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data))
summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+monitoring.binary+df16,data=train.data))

# No moderations were significant
```{r}
# Moderation: trauma * Network 17 --> PC1
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+trauma.binary*df18,data=train.data))

# Moderation: trauma * Network 15 --> PC1
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+trauma.binary*df16,data=train.data))

# Moderation: physical neglect * Network 17 --> PC1
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+monitoring.binary*df18,data=train.data))

# Moderation: physical neglect * Network 15 --> PC1
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+monitoring.binary*df16,data=train.data))

# Moderation: physical neglect * Network 17 --> PC3
summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+monitoring.binary*df18,data=train.data))

# Moderation: physical neglect * Network 15 --> PC3 
summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+monitoring.binary*df16,data=train.data))

# Moderation: sex * trauma * Network 17 --> PC1
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+trauma.binary*df18*sex,data=train.data))

# Moderation: sex * trauma * Network 15 --> PC1
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+trauma.binary*df16*sex,data=train.data))

# Moderation: sex * physical neglect * Network 17 --> PC1
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+monitoring.binary*df18*sex,data=train.data))

# Moderation: sex * physical neglect * Network 15 --> PC1
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+monitoring.binary*df16*sex,data=train.data))

# Moderation: sex * physical neglect * Network 17 --> PC3
summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+monitoring.binary*df18*sex,data=train.data))

# Moderation: sex * physical neglect * Network 15 --> PC3 
summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+monitoring.binary*df16*sex,data=train.data))
```

# Analyses of total FPN
```{r}
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+totalFPN,data=train.data)) #**
summary(lmer(thompson_PC2~interview_age+sex+(1|rel_family_id)+totalFPN,data=train.data)) #ns
summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+totalFPN,data=train.data)) #ns

summary(lmer(totalFPN~interview_age+sex+(1|rel_family_id)+monitoring.binary,data=train.data)) #ns
summary(lmer(totalFPN~interview_age+sex+(1|rel_family_id)+acceptance.binary,data=train.data)) #ns
summary(lmer(totalFPN~interview_age+sex+(1|rel_family_id)+abuse.binary,data=train.data)) #ns
summary(lmer(totalFPN~interview_age+sex+(1|rel_family_id)+trauma.binary,data=train.data)) #ns

summary(lmer(totalFPN~interview_age+sex+(1|rel_family_id)+monitoring.binary*sex,data=train.data)) #ns
summary(lmer(totalFPN~interview_age+sex+(1|rel_family_id)+acceptance.binary*sex,data=train.data)) #ns
summary(lmer(totalFPN~interview_age+sex+(1|rel_family_id)+abuse.binary*sex,data=train.data)) #ns
summary(lmer(totalFPN~interview_age+sex+(1|rel_family_id)+trauma.binary*sex,data=train.data)) #ns

# Port data over to matlab (womp) so I can run the Canlab multilevel mediation toolbox 
```{r}
write.csv(train.data,file="trainData_forMatlab.csv")
write.csv(test.data,file="testData_forMatlab.csv")
```

# Create a correlation matrix of all the important variables
```{r}
important.variables <- train.data[,c("interview_age","monitoring","acceptance","abuse","trauma","df2","df3","df4","df5","df6","df7","df8","df9","df10","df11","df12","df13","df14","df15","df16","df17","df18","thompson_PC1","thompson_PC2","thompson_PC3")]

colnames(important.variables)<-c("Age","Phys Neglect","Emo Neglect","Abuse","Trauma","PFN 1","PFN 2","PFN 3","PFN 4","PFN 5","PFN 6","PFN 7","PFN 8","PFN 9","PFN 10","PFN 11","PFN 12","PFN 13","PFN 14","PFN 15","PFN 16","PFN 17","General Cog (PC1)","Exec Fxn (PC2)","Learn/Mem (PC3)")

png("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/ABCD_PFNs_ACES_Cog_CorrMatrix.png",
    width = 8000, 
    height = 4000,
    res=300)
corrplot(cor(important.variables,use="complete.obs"),method=c("number"), tl.col = "black", col=brewer.pal(n=8, name="Spectral"))
dev.off() 
```
